The Concurrent Calculi Formalisation Benchmark

### 2024-02-23 short-paper/short-paper.tex
Minor edits and remove a comment before initial submit.

\section{Introduction}
The POPLMark challenge~\cite{POPLMark}
played a pivotal role in advancing the field of proof assistants,
libraries, and best practices for the mechanisation of programming language research. By providing a shared framework for systematically
evaluating mechanisation techniques, it catalysed a significant
shift towards publications that include mechanised proofs within the programming language research community.
POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme, targeting proofs using logical relations.
These initiatives had a narrow focus, and their authors recognised the importance of addressing topics such as coinduction and linearity in the future.

@@

Given those peculiarities, the syntax and semantics of $\pi$-calculi
have been mechanised from an early age (see~\cite{Melham1994}) with
many proof assistants and in many encoding styles.  Despite this,
almost all of these mechanisations rely on ad-hoc solutions to encode
scope extrusion.  They range from concrete encodings based on named
syntax~\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997,
  Perera2018} and locally-nameless representation~\cite{Castro2020}.
Nominal approaches are also common (see~\cite{Bengtson2009}), but they
may be problematic in proof assistants based on constructive type
theories.  An overall comparison is still lacking, but the case
study~\cite{AmbalLS21} develops four approaches to encoding binders in
Coq % as a first step towards developing tools for working with
in the context of higher-order process calculi.  The authors report that working directly
with de~Bruijn indices was easiest since the existing approaches
developed for $\lambda$-calculus binders worked poorly with scope
extrusion.

@@

The setting for our challenge is an untyped \(\pi\)-calculus augmented
with process replication in order to enable infinite behaviours.  We
do not include name passing since it is orthogonal to our aim of
exploring coinductive proof techniques.  We base our definition of
bisimilarity on a labelled transition system semantics and an
observability predicate describing the communication steps available
to a process.  The description of strong barbed bisimulation is one of
the first steps when studying the behaviour of process calculi, both
in textbooks (e.g.,~\cite{picalcbook}) and in existing mechanisations.
Though weak barbed congruence is a more common behavioural
equivalence, we prefer strong equivalences to simplify the theory by
avoiding the need to abstract over the number of internal transitions
in a trace.

@@

Combining techniques for mechanising the fundamental aspects covered in our benchmark is often not trivial.
While we have chosen to focus on the aspects individually to simplify the challenges, we are of course also interested in exploring how techniques interact.
Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}

>>>>

\section{Introduction}
The POPLMark challenge~\cite{POPLMark}
played a pivotal role in advancing the field of proof assistants,
libraries, and best practices for the mechanisation of programming language research. By providing a shared framework for systematically
evaluating mechanisation techniques, it catalysed a significant
shift towards publications that include mechanised proofs within the programming language research community.
POPLMark Reloaded~\cite{POPLMarkReloaded} introduced a similar programme for proofs using logical relations.
These initiatives had a narrow focus, and their authors recognised the importance of addressing topics such as coinduction and linearity in the future.

@@

Given those peculiarities, the syntax and semantics of $\pi$-calculi
have been mechanised from an early age (see~\cite{Melham1994}) with
many proof assistants and in many encoding styles.  Despite this,
almost all of these mechanisations rely on ad-hoc solutions to encode
scope extrusion.  They range from concrete encodings based on named
syntax~\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997,
  Perera2018} and locally-nameless representation~\cite{Castro2020}.
Nominal approaches are also common (see~\cite{Bengtson2009}), but they
may be problematic in proof assistants based on constructive type
theories.  An overall comparison is still lacking, but the case
study~\cite{AmbalLS21} explores four approaches to encoding binders in
Coq in the context of higher-order process calculi.  The authors report that working directly
with de~Bruijn indices was easiest since the approaches
developed for $\lambda$-calculus binders worked poorly with scope
extrusion.

@@

The setting for our challenge is an untyped \(\pi\)-calculus augmented
with process replication in order to enable infinite behaviours.  We
do not include name passing since it is orthogonal to our aim of
exploring coinductive proof techniques.  We base our definition of
bisimilarity on a labelled transition system semantics and an
observability predicate describing the communication steps available
to a process.  The description of strong barbed bisimulation is one of
the first steps when studying the behaviour of process calculi, both
in textbooks (\eg~\cite{picalcbook}) and in existing mechanisations.
Though weak barbed congruence is a more common behavioural
equivalence, we prefer strong equivalences to simplify the theory by
avoiding the need to abstract over the number of internal transitions
in a trace.

@@

Combining techniques for mechanising the fundamental aspects covered in our benchmark is often not trivial.
While we have chosen to focus on the aspects individually to simplify the challenges, we are of course also interested in exploring how techniques interact.
%Some authors have already described the difficulties that combining techniques may induce in a mechanisation, \eg~\cite{DBLP:conf/forte/ZalakainD21}.\todo{Add more references here?}
